Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-16 arxiv/proof_same_cov.tex
sec 3 wip

	\section{Proof of Lemma \ref{lem_error_same_cov}}\label{app_proof_error_same_cov}
	Next, we show that the concentration errors of $g_0(\cW), g_1(\cW), g_2(\cW)$ are all lower order terms compared to $\ex{g(\cW)}$.

>>>>

	\section{Missing Proofs of Theorem \ref{thm_many_tasks}}\label{app_proof_error_same_cov}

	\paragraph{Proof of equation \eqref{eq_gA}.}
	To facilitate the analysis, we introduce the following notations.
	Let
		\[ \cE \define {\sum_{k=1}^t \varepsilon^{(k)} A_k^{\top}}, \text{ and } \cW \define U_X U_X^{\top} \cE (AA^{\top})^{-1}. \]
	For any $j = 1,2,\dots, t$, let
	\begin{align*}
		H_j &\define B^{\star} A^{\top} (AA^{\top})^{-1} A_j - \beta^{(j)}, \\
		E_j &\define \cW A_j - \varepsilon^{(j)}.
	\end{align*}
	Then, we have that
	\[ g(A) = \sum_{j=1}^t \bignorm{H_j + E_j}^2. \]
	To compute the expectation of $g(A)$, we first observe that for any first order terms of $\varepsilon^{(1)}, \dots, \varepsilon^{(t)}$, their expectation is zero. Therefore,
	\begin{align*}
		\exarg{\set{\varepsilon^{(j)}}_{j=1}^t }{g(A)} = \sum_{j=1}^t \norm{H_j}^2 + \exarg{\set{\varepsilon^{(j)}}_{j=1}^t, X}{\sum_{j=1}^t \norm{E_j}^2}.
	\end{align*}
	Hence, it suffices to show that the second part of the equation above is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	Conditional on $X$, we have that
	\begin{align*}
		\exarg{\set{\varepsilon^{(j)}}_{j=1}^t}{\norm{E_j}^2}
		= \exarg{\set{\varepsilon^{(j)}}_{j=1}^t}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
	\end{align*}
	We consider both terms in the above equation one by one.
	First, note that
	\begin{align}
		\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE (AA^{\top})^{-1} \cE^{\top}}. \label{eq_proof_same_cov_1}
	\end{align}
	We observe that
	\begin{align*}
			\exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\cE (AA^{\top})^{-1} \cE^{\top}}
		&= \exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above, the first equation uses the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The second equation uses the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{-1} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
	\begin{align*}
		\sum_{j=1}^t \norm{\cW A_j}^2 = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
	\end{align*}
	because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	Next, note that
	\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW \cE^{\top}}, \]
	which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$. Hence the proof is complete.

	\paragraph{Proof of Lemma \ref{lem_error_same_cov}.} Next, we show that the concentration errors of $g_0(\cW), g_1(\cW), g_2(\cW)$ are all lower order terms compared to $\ex{g(\cW)}$.

### 2020-09-16 arxiv/setup.tex
sec 3 wip

\subsection{Notations.}
We shall say an event $\Xi$ holds with high probability (w.h.p.) if for any fixed $D>0$, $\P(\Xi)\ge 1- p^{-D}$ for large enough $p$. Moreover, we shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p\to \infty$.
For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.
We shall refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices following the standard notations in high-dimensional statistics.
%\subsection{Bias and Variance}

@@

We introduce several standard assumptions on the random matrices we study.
\begin{assumption}\label{assume_rm}
	Let $c > 4$ be a fixed value.
	Let $X = Z \Sigma^{1/2} \in\real^{n\times p}$ be a random matrix where $Z \in \real^{n\times p}$ consists of i.i.d. entries with zero mean and unit variance and $\Sigma \in \real^{p\times p}$ is a positive semidefinite matrix.
	We assume that $\Sigma$ and $Z$ satisfy the following (standard) conditions, respectively.
	\begin{enumerate}
		\item[i)] Bounded moments: for every entry of $Z$, we assume that its $c$-th moment exists, that is, there exist a fixed constant $C > 0$ such that
			\begin{equation}\label{assmAhigh}
				\ex{\abs{Z_{i,j}}^{c}} \le C, \text{ for any } 1\le i \le n, 1\le j \le p.
			\end{equation}
		\item[ii)] Bounded singular values: the singular values of $\Sigma$ are bounded within a fixed range, that is, there exists two fixed values $C > C' > 0$ such that all of $\Sigma$'s singular values are bounded within $C'$ and $C$.
	\end{enumerate}
\end{assumption}

\begin{lemma}[cf. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
	Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	Let $\e$ be any fixed value within $(0, \frac{c - 4}{2c})$.
	With high probability over the randomness of $X$, we have that
	\[ \bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{\sigma^2 \cdot p}{n - p} + \OO(p^{-\e}). \]
\end{lemma}

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise of every task, we assume that all $l$, the $l$-th moment of the noise entries exists:
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{k, j}}^l \le C}, \text{ for all } j = 1, \dots, n_k.
\end{align}

>>>>

\subsection{Notations}
We shall say an event $\Xi$ holds with high probability (w.h.p.) if for any fixed $D>0$, $\P(\Xi)\ge 1- p^{-D}$ for large enough $p$. Moreover, we shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p\to \infty$.
For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.
We shall refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices following the standard notations in high-dimensional statistics.
%\subsection{Bias and Variance}

@@

We introduce several standard assumptions on the random matrices we study.
\begin{assumption}\label{assume_rm}
	Let $c > 4$ be a fixed value.
	Let $X = Z \Sigma^{1/2} \in\real^{n\times p}$ be a random matrix where $Z \in \real^{n\times p}$ consists of i.i.d. entries with zero mean and unit variance and $\Sigma \in \real^{p\times p}$ is a positive semidefinite matrix.
	We assume that $\Sigma$ and $Z$ satisfy the following (standard) conditions, respectively.
	\begin{enumerate}
		\item Bounded moments: for every entry of $Z$, we assume that its $c$-th moment exists, that is, there exist a fixed constant $C > 0$ such that
			\begin{align}\label{assmAhigh}
				\ex{\abs{Z_{i,j}}^{c}} \le C, \text{ for any } 1\le i \le n \text{ and } 1\le j \le p.
			\end{align}
		\item Bounded singular values: the singular values of $\Sigma$ are bounded within a fixed range, that is, there exists two fixed values $C_{\max} > C_{\min} > 0$ such that all of $\Sigma$'s singular values are bounded within $C_{\min}$ and $C_{\max}$.
	\end{enumerate}
\end{assumption}

Under the above assumption, we state several well-known facts regarding the random matrix $X$.

\begin{fact}\label{lem_minv}
	Let $\e$ be any fixed value within $(0, \frac{c - 4}{2c})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item The sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-\e})$, cf. Theorem 2.4 in \citet{isotropic}.
	\end{enumerate}
\end{fact}

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise component, we assume that all of its moments exist.
More precisely, there exists a fixed function $C(\cdot) : \mathbb{Z} \rightarrow \real^+$ such that for any $a = 1, 2, \dots, \infty$, we have that
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{j}^{(i)}}^a} \le C(a), \text{ for any } 1\le i\le t \text{ and } 1\le j\le n_i.
\end{align}
